---
title: "Kovach-Townsend Spectral Processing Workflow"
output: html_document
date: "2023-10-12"
editor_options: 
  chunk_output_type: console
runtime: shiny
---

## Define Folders

##### Define Spectral Folder
This script will recurse through a folder and all subfolders to select all spectral files.  In the final output table, each spectral measurement will be a row, the first column will be the file name, and all subsequent columns will be spectral wavelengths.

```{r Define Folders}

specfolder="Demo/Spectra_Test"

```

##### Define Models Folder
If applying partial least squares regression models (PLSR) to predict traits on the spectral data, the location of those models will be defined here.  For existing Townsend PLSR models, the leaf measured material informs the folder selection.  In the GitHub repository, within the "Models" folder, there are two folder named "models_ASD_fresh_spectra" and "models_ASD_dry_spectra."  We choose the models that match the sample preparation type.  The script will recurse through the folder and apply all models in the folder.

```{r Define Models}

specmodelpath="Models/models_ASD_fresh_spectra"

```

## Setup and Initialize
Here we set the packages necessary to do the processing of spectra from raw spectral files.  This script will handle any of the three common instruments (ASD, Spectral Evolution, and Spectral Vista).  It will load and carry out jump correction, and outputs both original and vector normalized spectra, as well as resampling to 5nm.  It will also apply PLSR models to spectra.  Existing models are for plant foliar material, though any PLSR model set can be used for other applications.

```{r Initialize, echo=T, results='hide', message=F}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("doParallel",repos = "http://cran.us.r-project.org")
# install.packages("spectrolab",repos = "http://cran.us.r-project.org")
# install.packages("dplyr",repos = "http://cran.us.r-project.org")
# install.packages("tidyr",repos = "http://cran.us.r-project.org")
# install.packages("ggplot2",repos = "http://cran.us.r-project.org")
# install.packages("reshape2",repos = "http://cran.us.r-project.org")
# install.packages("stringr",repos = "http://cran.us.r-project.org")
# install.packages("tools",repos = "http://cran.us.r-project.org")
# install.packages("cowplot",repos = "http://cran.us.r-project.org")
# install.packages("GGally",repos = "http://cran.us.r-project.org")

require(doParallel)
require(spectrolab)
require(dplyr)
require(tidyr)
require(ggplot2)
require(cowplot)
require(reshape2)
require(stringr)
require(tools)
require(GGally)

set.seed(112345)
cores=detectCores()-1

split_path=function(path) {
  if (dirname(path) %in% c("_", path)) return(basename(path))
  return(c(basename(path),split_path(dirname(path))))
}

cn=function(x)
{
  print(as.matrix(colnames(x)))
}
```

## Import Spectra
The following code parallel processes the spectral files to import them from the defined folder and carry out a jump correction specific for the sensor data, followed by a normalization, and then a resampling to 5nm.  All four data types:
- raw spectra
- jump corrected
- jump corrected and unit vector normalized
- jump corrected, unit vector normalized, and resampled to 5nm will be available, though the model application will be with the vector normalized data (as the models were built using vector normalized data).

```{r Import Spectra,message=F}

#----Import Spectra, Jump Correct, Vector Normalize, Resample----
for (j in c("asd","sed","sig"))
{
  if (exists('finaloutput')){rm(finaloutput)}
  folderlist=unique(sub("/[^/]+$","",list.files(path=specfolder,
                                                pattern=paste0(".",j,"$"),
                                                recursive=TRUE,
                                                full.names = TRUE)))
  if (length(folderlist)!=0){
    cl=makeCluster(cores)
    registerDoParallel(cl)
    
    finaloutput=foreach(i=folderlist,
                        .combine=rbind,
                        .packages=c('spectrolab'))%dopar%
      {
        spectra=read_spectra(path=i,format=j)
        data.frame(spectra,folder=i,format=j,check.names=FALSE)
        };FST=as_spectra(finaloutput[,1:2152],name_idx=1)
    
    stopCluster(cl)
    
    if (j=="asd"||j=="sig"){
      splice_val=ifelse(j=="asd",c(1000, 1800),c(990, 1900))
      FST=match_sensors(FST,splice_at=splice_val)
    };FSTdf=data.frame(FST,check.names=FALSE)
    if (!exists('FFO')){FFO=FSTdf[0,]}
    FFO=rbind(FFO,FSTdf)
  }};FSTfinal=as_spectra(FFO,name_idx=1);FSTfinal_norm=normalize(FSTfinal);FSTfinal_sd=sd(FSTfinal)

FST_f=data.frame(FSTfinal)
FST_n=data.frame(FSTfinal_norm)

head(FST_f)[c(1,402:412)] #Original Spectra

#----Process Spectra, Vector Normalize, and Resample----
headcount=ncol(FST_f)-2151
specstart=headcount+1

cl=makeCluster(cores);registerDoParallel(cl);sampledata_VN=foreach(i=1:nrow(FST_f),
                                                                   .combine=rbind)%dopar%
  {
    a=FST_f[i,]
    b=sqrt(rowSums(a[,specstart:ncol(a)]^2))
    d=a[,-c(1:headcount)]
    sample_name=a[,c(1:headcount)]
    f=d/b
    cbind(sample_name,f)
  }
stopCluster(cl)

head(sampledata_VN)[c(1,402:412)] #Vector Normalized Spectra

sampledata=sampledata_VN

sampledata_wav=sampledata[,-c(1:headcount)]
sampledata_head=sampledata[,c(1:headcount)]
sampledata_5nm_wav=sampledata_wav[,seq_len(ncol(sampledata_wav)) %% 5 == 1]
sampledata_5nm=cbind(sampledata_head,sampledata_5nm_wav)
sampledata=sampledata_5nm

head(sampledata)[c(1,82:92)] #Vector Normalized Resampled Spectra

```

## Plots and Outlier Detection
Following are some methods to plot the data you've extracted.  Some are built using the Spectrolab package, while others are written from existing R code.  Using individual wavelengths is a good way to quickly assess if there are any significant outliers that should be removed.

```{r Plots}

plot_interactive(FSTfinal,
                 colpalette = function(n) RColorBrewer::brewer.pal(n, "Dark2"),
)

plot(FSTfinal, ylab = "value", xlab = "band", lty = 1, type = "l")

plot_regions(
      FSTfinal,
      regions = default_spec_regions(),
      col = grDevices::rgb(0.7, 0.7, 0.7, 0.3),
      border = FALSE,
      add = TRUE,
      add_label = TRUE,
      cex_label = 1,
)

plot(FSTfinal_norm, ylab = "value", xlab = "band", lty = 1, type = "l")

plot_regions(
      FSTfinal,
      regions = default_spec_regions(),
      col = grDevices::rgb(0.7, 0.7, 0.7, 0.3),
      border = FALSE,
      add = TRUE,
      add_label = TRUE,
      cex_label = 1,
)

matplottable=FST_f[c(-1)]
matplot(t(matplottable), type = 'l',xaxt="n",xlab="Wavelength (nm)",ylab="Reflectance (%)")
axis(1, at=c(seq(from=0,to=2150,by=50)), labels=c(seq(from=350,to=2500,by=50)))

#Outlier Detection
minmin=which.min(FST_f$X1200)
min(FST_f$X1200,na.rm=T)
FST_f[minmin,1]

#Outlier Detection
maxmax=which.max(FST_f$X550)
max(FST_f$X550,na.rm=T)
FST_f[maxmax,1]

# #Correlation Plot
# ggcorr(sampledata[c(-1)])

```

## Grouped Outlier Detection
This is a grouped manual outlier detection for samples with replicates.  It vector normalizes the spectra, and uses a sample and replicate value (extracted in this case from the file nomenclature) to display them separately.

```{r Grouped Outlier Detection, warnings=F,message=F}

sample_data_plot=FST_f

sample_data_plot$Sample=str_split_fixed(sample_data_plot$sample_name,"_",n=4)[,1]
sample_data_plot$Replicate=str_split_fixed(sample_data_plot$sample_name,"_",n=4)[,2]

sample_data_plot_sub=sample_data_plot[,c(ncol(sample_data_plot)-1,ncol(sample_data_plot),4:ncol(sample_data_plot)-2)]

tic=0;for (i in unique(sample_data_plot_sub$Sample)){
  a=subset(sample_data_plot_sub,sample_data_plot_sub$Sample==i)
  aa=a[,-c(1:2)]
  b=sqrt(apply(aa[,1:2151]^2,1,sum))
  d=aa[,1:2151]/b
  e=cbind(a[,1:2],d)
  f=melt(e, id.vars=c("Sample","Replicate"))
  f$Replicate=as.factor(f$Replicate)
  print(ggplot(f, aes(x=as.numeric(variable), y=value, group=Replicate,color=Replicate))+
    geom_line()+
    ggtitle(paste("Sample:",f$Sample[1]))+
    ylab("value")+
    theme_bw()+
    scale_x_continuous(name="band",
                       limits=c(0, 2151),
                       breaks=seq(150, 2151,by=500),
                       labels=seq(500,2500, by=500)))
  tic=tic+1;print(paste0(tic,"/",length(unique(sample_data_plot_sub$Sample))))
  readline(prompt="Press [enter] to continue")
};rm(a,b,d,e,f,i,sample_data_plot_sub,tic)

```


## Predict Traits
This code block will predict traits using the models specified in the folder at the beginning of the script.  It will iterate over all the models in that folder with all the spectra extracted above.  Output here is a table with the sample name as the first column, and all additional columns being the traits predicted, and the uncertainties associated with them.

```{r Predict Traits, warnings=F}

#----Predict Traits----

modeldirectory=list.files(path=specmodelpath,
                          pattern=".csv$",
                          recursive=FALSE,
                          full.names = TRUE)

cl=makeCluster(cores);registerDoParallel(cl);finished_output=foreach(model=modeldirectory,
                                                                     .combine=rbind)%do%
  {
    modelname=split_path(model)[1]
    modelcoeffs_mat=data.matrix(read.csv(model, header = TRUE, row.names = 1,sep = ","))
    matplottable=modelcoeffs_mat[1,c(12:412)]
    matplot(matplottable, main=modelname, type = 'l', xaxt="n", xlab="Wavelength (nm)", ylab="Coefficient")
    axis(1, at=c(seq(from=0,to=400,by=2)), labels=c(seq(from=400,to=2400,by=10)))
    sampledata_tail=sampledata[,-c(1:headcount)]
    sampledata_head=sampledata[,c(1:headcount)]
    sampledata_tail$constant=1
    sampledata_tail=cbind(intercept=sampledata_tail[,ncol(sampledata_tail)],sampledata_tail[,c(1:ncol(sampledata_tail)-1)])
    sampledata_tail_mat=as.matrix(sampledata_tail)
    modeloutput=sampledata_tail_mat %*% t(modelcoeffs_mat)
    modeloutput_bind=cbind(sampledata_head,as.data.frame(modeloutput))
    data.frame(modelname=modelname,
               sample_name=sampledata_head,
               t_mean=rowMeans(modeloutput),
               t_std=apply(modeloutput,1,sd),
               as.data.frame(modeloutput),
               check.names=FALSE)
  };stopCluster(cl)

predplot=finished_output[,c(1,3:4)]
predplot$Group=str_split_fixed(finished_output$sample_name,"_",n=2)[,1]

predsub=finished_output[,c(1:4)]

predspread=pivot_wider(data=predsub,id_cols=sample_name,names_from=modelname,values_from=c("t_mean","t_std"))
predunlist=as.data.frame(unnest(predspread))
colnames(predunlist)=sub(".csv", "", colnames(predunlist))
# write.csv(predunlist,"SpectroPredictR_finished_model_output.csv",row.names=FALSE)
predunlist[1:4,]

```

## Plot Traits
To better visualize trait distributions, this script will produce box plots with a "grouping" based on the sample name.  It plots both the mean trait for that group, and the uncertainties associated with it.  The group can be redifined to any column that uniquely defines the data.  An example would be grouping by growth form which would give us the distribution over that entire group.

```{r Plot Traits}
##### Plot Boxplots #####
for (i in unique(predplot$modelname)){

  datanow=subset(predplot,modelname==i)

  d=ggplot(datanow,
           aes(x=Group,y=t_mean))+
    ggtitle(paste0(file_path_sans_ext(i)," (Fresh Samples) [n=",nrow(datanow),"]"))+
    geom_boxplot(color="blue",fill="lightblue",size=0.5)+
    coord_flip()+
    xlab("Sample")+
    ylab("Mean of Predictions")+
    scale_color_brewer(palette="Set1")+
    theme(text = element_text(size=15))+
    scale_x_discrete(limits=rev(levels(as.factor(datanow$Group))))

  e=ggplot(datanow,
           aes(x=Group,y=t_std,))+
    ggtitle("    ")+
    geom_boxplot(color="black",fill="yellow",size=0.5)+
    ylab("StDev of Predictions")+
    scale_color_brewer(palette="Set1")+
    coord_flip()+
    scale_x_discrete(limits=rev(levels(as.factor(datanow$Group))))+
    theme(axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
          legend.position="none")

  print(plot_grid(d,e,rel_widths = c(4,1)))

  # ggsave(
  #   file=paste0("Plots/",file_path_sans_ext(i),"_dry_plot.png"),
  #   width = 11,
  #   height = 8.5,
  #   units="in",
  #   dpi = 300
  # )
}
```

